Measuring security performance in the Android ecosystem

### 2015-06-05 ccs.tex
Fix some todos
Adjust display hacks, less sloppy, more interfootnotelinepenalty

@@

\section{Comparison with other data}\label{sec:representative}
Here we show that the \da\ data~\cite{Wagner2013} is representative for our purposes by comparing it with two other sources of data.
The \da\ data is collected from devices across the world where users have voluntarily installed the app.
We have obtained comparable data on 5\,290 devices from a multinational FTSE 100 company's mobile device management database which includes company and employee owned devices, and from 5\,170\,000 matching User-Agent headers on all HTTP traffic for 30\% of Rwanda for a week.
\todo{citation for Rwanda dataset}
\begin{figure}
\centering
\includegraphics[width=\columnwidth]{figures/dists}
\caption{Comparison between FTSE, User-Agent and the corresponding \da\ data, error bars indicate 95\% confidence intervals.}
\label{fig:dists}
\end{figure}
We used the data from the FTSE 100 company for the week ending 2015-04-27 and the User-Agent data was collected between 2015-01-31 and 2015-02-12.
Figure~\ref{fig:dists} shows the proportion of devices running each Android OS version in the two comparison data sets and the comparable periods from \da.
The general pattern this shows is that in the FTSE data newer versions are more popular than in the \da\ data and that in the Rwanda data old versions are more popular than in the \da\ data.
\todo{Can we quantify this?}

@@

\interfootnotelinepenalty=10000 %To stop the lifehacker footnote below from spanning multiple pages and breaking
We use User-Agent strings from Rwandan internet traffic to examine the version distribution of Android OS versions.
User-Agent strings have been used to investigate the timeliness of web browser updates, with at most 80\% of Firefox users running the most recent version~\cite{Frei2008}.
The same analysis was used to show that Chrome's use of silent updates seems to increase uptake of upgrades~\cite{Duebendorfer2010} with 97\% of users running the latest version within 3 weeks of release.
By way of comparison, Android's update process is manual.
The user is notified an update exists, but further action is required, including downloading the update and rebooting the phone to enable installation.
The phone must have sufficient charge to perform the update and the device itself is rendered inoperable during the update process, two factors which might prevent or delay the update process from taking place.
In our data we are unable to determine why a device is not updated. 
It is possible that many updates arrive at handsets, but are simply not installed.
Anecdotal evidence at least suggests that it is the lack of updates rather the lack of installation which is the major problem at present and \S\ref{sec:security_scoring:results} supports this. Further work is required to tease these numbers apart.
Partly this is the result of the fact that an operating system update is being installed and so a reboot is required, but Chrome installs the new version side by side with the old one and switches the next time it is restarted.
The same technique would be more difficult on phones with limited storage space (as many cheap Android phones have barely enough space to install just the update) but is a plausible improvement for more high-end devices.
Google is deploying the same silent update technique through Google Play Services\footnote{\href{http://lifehacker.com/why-google-play-services-are-now-more-important-than-an-975970197}{http://lifehacker.com/why-google-play-services-are-now-more-important-than-an-975970197}} which automatically installs updates for core Google components of Android, this also bypasses the device manufacturer and network operator.

@@

\sloppy
\section*{Acknowledgements}
\identifying{
This work was supported by a Google focussed research award; and the EPSRC [grant number EP/P505445/1].
Thanks to Sherif Akoush and Ripduman Sohan for supplying the Rwanda data and to staff at the FTSE 100 company for supplying their mobile device management data.
}

\printbibliography

>>>>

% To stop problems with links in footnotes spanning columns or pages.
\interfootnotelinepenalty=10000

@@

\section{Comparison with other data}\label{sec:representative}
Here we show that the \da\ data~\cite{Wagner2013} is representative for our purposes by comparing it with two other sources of data.
The \da\ data is collected from devices across the world where users have voluntarily installed the app.
We have obtained comparable data on 5\,290 devices from a multinational FTSE 100 company's mobile device management database which includes company and employee owned devices, and from 5\,170\,000 matching User-Agent headers on all HTTP traffic for 30\% of Rwanda for a week.
\todo{citation for Rwanda dataset}
\begin{figure}
\centering
\includegraphics[width=\columnwidth]{figures/dists}
\caption{Comparison between FTSE, User-Agent and the corresponding \da\ data, error bars indicate 95\% confidence intervals.}
\label{fig:dists}
\end{figure}
We used the data from the FTSE 100 company for the week ending 2015-04-27 and the User-Agent data was collected between 2015-01-31 and 2015-02-12.
Figure~\ref{fig:dists} shows the proportion of devices running each Android OS version in the two comparison data sets and the comparable periods from \da.
The general pattern this shows is that in the FTSE data newer versions are more popular than in the \da\ data and that in the Rwanda data old versions are more popular than in the \da\ data.
\todolater{Can we quantify this?}

@@

We use User-Agent strings from Rwandan internet traffic to examine the version distribution of Android OS versions.
User-Agent strings have been used to investigate the timeliness of web browser updates, with at most 80\% of Firefox users running the most recent version~\cite{Frei2008}.
The same analysis was used to show that Chrome's use of silent updates seems to increase uptake of upgrades~\cite{Duebendorfer2010} with 97\% of users running the latest version within 3 weeks of release.
By way of comparison, Android's update process is manual.
The user is notified an update exists, but further action is required, including downloading the update and rebooting the phone to enable installation.
The phone must have sufficient charge to perform the update and the device itself is rendered inoperable during the update process, two factors which might prevent or delay the update process from taking place.
In our data we are unable to determine why a device is not updated. 
It is possible that many updates arrive at handsets, but are simply not installed.
Anecdotal evidence at least suggests that it is the lack of updates rather the lack of installation which is the major problem at present and \S\ref{sec:security_scoring:results} supports this. Further work is required to tease these numbers apart.
Partly this is the result of the fact that an operating system update is being installed and so a reboot is required, but Chrome installs the new version side by side with the old one and switches the next time it is restarted.
The same technique would be more difficult on phones with limited storage space (as many cheap Android phones have barely enough space to install just the update) but is a plausible improvement for more high-end devices.
Google is deploying the same silent update technique through Google Play Services\footnote{\href{http://lifehacker.com/why-google-play-services-are-now-more-important-than-an-975970197}{http://lifehacker.com/why-google-play-services-are-now-more-important-than-an-975970197}} which automatically installs updates for core Google components of Android, this also bypasses the device manufacturer and network operator.

@@

\section*{Acknowledgements}
\identifying{
This work was supported by a Google focussed research award; and the EPSRC [grant number EP/P505445/1].
Thanks to Sherif Akoush and Ripduman Sohan for supplying the Rwanda data and to staff at the FTSE 100 company for supplying their mobile device management data.
}
\sloppy
\printbibliography
